A meta-analysis of mental rotation ability in the first years of life

### 2022-04-22 code/analysis.Rmd
Rename `dat` objects

    # Compute empirical correlation based on sd_z and condition SDs
    d_z = case_when(
      !is.na(d_z_t) ~ d_z_t,
      !is.na(d_z_f) ~ d_z_f,
      !is.na(d_z_diff) ~ d_z_diff
    ),
    sd_z = mean_diff / d_z,
    ri = (sd_z^2 - sd_novel^2 - sd_familiar^2) / (-2 * sd_novel * sd_familiar)
  ) -> dat

# Overview of the different effect sizes
dat %>%
  select(
    article,
    group,
    gender_split,
    gi,
    di,
    di_type,
    d,
    d_z_t,
    d_z_f,
    d_z_diff,
    d_av,
    d_nov_pref,
    ri
  ) %>%
  print(n = Inf)

# Get empirical estimate of the correlation between dependent samples
summary(dat$ri)

# Compute standard error of Cohen's d based on assumed correlation
# See Hedges' formula on p. 253 in http://dx.doi.org/10.20982/tqmp.14.4.p242
# This will need a sensitivity analysis regarding the values of `r_assumed`
r_assumed <- 0.5
dat %>%
  mutate(
    ni = sample_size,
    vi = (dfi / (dfi - 2)) * ((2 * (1 - r_assumed)) / ni) *
      (1 + gi^2 * (ni / (2 * (1 - r_assumed)))) - (gi^2 / ji^2),
    sei = sqrt(vi)
  ) %>%
  filter(!is.na(gi)) %>%
  select(
    article,
    group,
    experiment,
    gender_split,
    gi,
    ni,
    vi,
    sei,
    gender,
    age
  ) -> dat_r

# Show number of experiments per gender grouping strategy
table(dat_r$gender_split)

# Extract mixed/split experiments only
dat_mixed <- filter(dat_r, gender_split %in% c("mixed", "mixed_only"))
dat_split <- filter(dat_r, gender_split %in% c("split", "split_only"))

# Combine both strategies, preferring mixed experiments
dat_both_mixed <- filter(
  dat_r, gender_split %in% c("mixed", "mixed_only", "split_only")
)

# Combine both strategies, preferring mixed experiments
dat_both_split <- filter(
  dat_r, gender_split %in% c("split", "mixed_only", "split_only")
)
```

@@

```{r, compute_meta, echo=TRUE}
# Three-level model
res_ml <- rma.mv(
  gi, vi,
  random = ~ 1 | article / experiment,
  data = dat_both_mixed,
  slab = experiment
)
print(res_ml)
# forest(res_ml)

@@

# Apply sliding difference contrast coding for the gender categories
# This means that the Intercept will be the grand mean (i.e., the average across
# studies) whereas the betas will test female vs. mixed and male vs. female
(contrasts(dat_both_mixed$gender) <- MASS::contr.sdif(3))

# Meta-regression with gender
res_gender <- rma.mv(
  gi, vi,
  mods = ~ gender - 1,
  random = ~ 1 | article / experiment,
  data = dat_both_mixed,
  slab = experiment
)
print(res_gender)
# forest(res_gender)

@@

# Meta-regression with age
res_age <- rma.mv(
  gi, vi,
  mods = ~age,
  random = ~ 1 | article / experiment,
  data = dat_both_mixed,
  slab = experiment
)
print(res_age)
# forest(res_age)

# Meta-regression with gender, age, and their interaction
res_full <- rma.mv(
  gi, vi,
  mods = ~ gender * age,
  random = ~ 1 | article / experiment,
  data = dat_both_mixed,
  slab = experiment
)
print(res_full)
# forest(res_full)
```

@@

# Run prior predictive simulation
res_prior <- brm(
  gi | se(sei) ~ 1 + (1 | article / experiment),
  data = dat_both_mixed,
  prior = prior_c,
  sample_prior = "only",
  save_pars = save_pars(all = TRUE),
  chains = 4,
  iter = 2000,
  cores = 4,
  control = list(adapt_delta = 0.99)
)
summary(res_prior)
# plot(res_prior)

@@

```{r, fig2}
# Forest plot inputs
level <- 0.05
ndraws <- 1000
dat_plot <- dat_both_mixed
res_plot <- res_ml
res_plot_brm <- res_brm

@@

# Get draws from the Bayesian posterior distribution
draws <- epred_draws(res_brm, dat_plot, ndraws = ndraws)
meta_draws <- spread_draws(res_brm, b_Intercept, ndraws = ndraws) %>%
  mutate(article = "Three-level model", experiment = "zzz") # For the last line

# Create forest plot
dat_plot %>%
  # Compute frequentist confidence intervals
  mutate(
    ci_lb = gi - qnorm(level / 2, lower.tail = FALSE) * sqrt(vi),
    ci_ub = gi + qnorm(level / 2, lower.tail = FALSE) * sqrt(vi),
    fontface = "plain" # For study labels and confidence interval
  ) %>%
  # Add meta-analytic effect size + CI
  bind_rows(list(
    article = "Three-level model",
    group = "",
    experiment = "zzz", # Just so it's displayed as the last line
    gi = res_plot$b,
    ci_lb = res_plot$ci.lb,
    ci_ub = res_plot$ci.ub,
    fontface = "bold"
  )) %>%
  # Format effect size + CI for printing them on the plot
  mutate(ci_print = print_ci(gi, ci_lb, ci_ub)) %>%
  # Reverse order of experiments so they are plotted alphabetically
  mutate(experiment = forcats::fct_rev(factor(experiment))) %>%
  arrange(experiment) %>%
  # Show article labels only for the first experiment per article
  mutate(
    article = if_else(article == lead(article, default = ""), "", article)
  ) %>%
  # Prepare plot
  ggplot(aes(x = gi, y = experiment)) +
  geom_vline(xintercept = seq(-1, 2, 1), size = 0.8, color = "grey90") +
  geom_vline(xintercept = seq(-1.5, 2.5, 1), size = 0.4, color = "grey90") +
  # Add article and group labels as text to the left
  geom_text(aes(x = -5.6, label = article, fontface = fontface), hjust = 0) +
  geom_text(aes(x = -3.1, label = group, fontface = fontface), hjust = 0) +
  # Add experiment specific effect size + CI as text to the right
  geom_text(aes(x = 4.4, label = ci_print, fontface = fontface), hjust = 1) +
  # Add Bayesian highest posterior density interval in the background
  stat_interval(aes(x = .epred), draws, point_interval = "mean_hdi") +
  stat_interval(aes(x = b_Intercept), meta_draws, point_interval = "mean_hdi") +
  scale_color_grey(
    start = 0.9, end = 0.6,
    labels = as_mapper(~ scales::percent(as.numeric(.x)))
  ) +
  labs(color = "Bayesian\nHPDI") +
  # Add study specific effect size + CIs on top
  ggnewscale::new_scale_color() +
  geom_errorbar(aes(xmin = ci_lb, xmax = ci_ub, color = gender)) +
  geom_point(aes(size = ni, color = gender), shape = 15) + # Or (1 / vi)?
  scale_color_brewer(palette = "Dark2", na.translate = FALSE) +
  # Add meta-analytic effect size as a line
  annotate("segment", x = res_plot$b, xend = res_plot$b, y = 0.7, yend = 1.3) +
  annotate("errorbar", xmin = res_plot$ci.lb, xmax = res_plot$ci.ub, y = 1.0) +
  # Styling
  coord_cartesian(expand = FALSE) +
  scale_x_continuous(breaks = seq(-1.5, 2.5, 0.5)) +
  labs(
    x = expression("Hedges'" ~ italic("g")),
    size = "Sample\nsize",
    color = "Gender"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.title.x = element_text(hjust = 0.584),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    legend.title = element_text(hjust = 0.5),
    panel.grid = element_blank(),
    text = element_text(family = "Helvetica")
  )

>>>>

    # Compute empirical correlation based on sd_z and condition SDs
    d_z = case_when(
      !is.na(d_z_t) ~ d_z_t,
      !is.na(d_z_f) ~ d_z_f,
      !is.na(d_z_diff) ~ d_z_diff
    ),
    sd_z = mean_diff / d_z,
    ri = (sd_z^2 - sd_novel^2 - sd_familiar^2) / (-2 * sd_novel * sd_familiar)
  ) -> dat_all

# Overview of the different effect sizes
dat_all %>%
  select(
    article,
    group,
    gender_split,
    gi,
    di,
    di_type,
    d,
    d_z_t,
    d_z_f,
    d_z_diff,
    d_av,
    d_nov_pref,
    ri
  ) %>%
  print(n = Inf)

# Get empirical estimate of the correlation between dependent samples
summary(dat_all$ri)

# Compute standard error of Cohen's d based on assumed correlation
# See Hedges' formula on p. 253 in http://dx.doi.org/10.20982/tqmp.14.4.p242
# This will need a sensitivity analysis regarding the values of `r_assumed`
r_assumed <- 0.5
dat_all %>%
  mutate(
    ni = sample_size,
    vi = (dfi / (dfi - 2)) * ((2 * (1 - r_assumed)) / ni) *
      (1 + gi^2 * (ni / (2 * (1 - r_assumed)))) - (gi^2 / ji^2),
    sei = sqrt(vi)
  ) %>%
  filter(!is.na(gi)) %>%
  select(
    article,
    group,
    experiment,
    gender_split,
    gi,
    ni,
    vi,
    sei,
    gender,
    age
  ) -> dat_all_r

# Show number of experiments per gender grouping strategy
table(dat_all_r$gender_split)

# Extract split or mixed experiments only
dat_split <- filter(dat_all_r, gender_split %in% c("split", "split_only"))
dat_mixed <- filter(dat_all_r, gender_split %in% c("mixed", "mixed_only"))

# Combine both strategies, preferring split experiments
dat_both_split <- filter(
  dat_all_r, gender_split %in% c("split", "mixed_only", "split_only")
)

# Combine both strategies, preferring mixed experiments
dat_both_mixed <- filter(
  dat_all_r, gender_split %in% c("mixed", "mixed_only", "split_only")
)

# We go ahead with the last solution as to use the maximum amount of information
# while not biasing the gender results
dat <- dat_both_mixed
```

@@

```{r, compute_meta, echo=TRUE}
# Three-level model
res_ml <- rma.mv(
  gi, vi,
  random = ~ 1 | article / experiment,
  data = dat,
  slab = experiment
)
print(res_ml)
# forest(res_ml)

@@

# Apply sliding difference contrast coding for the gender categories
# This means that the Intercept will be the grand mean (i.e., the average across
# studies) whereas the betas will test female vs. mixed and male vs. female
(contrasts(dat$gender) <- MASS::contr.sdif(3))

# Meta-regression with gender
res_gender <- rma.mv(
  gi, vi,
  mods = ~ gender - 1,
  random = ~ 1 | article / experiment,
  data = dat,
  slab = experiment
)
print(res_gender)
# forest(res_gender)

@@

# Meta-regression with age
res_age <- rma.mv(
  gi, vi,
  mods = ~age,
  random = ~ 1 | article / experiment,
  data = dat,
  slab = experiment
)
print(res_age)
# forest(res_age)

# Meta-regression with gender, age, and their interaction
res_full <- rma.mv(
  gi, vi,
  mods = ~ gender * age,
  random = ~ 1 | article / experiment,
  data = dat,
  slab = experiment
)
print(res_full)
# forest(res_full)
```

@@

# Run prior predictive simulation
res_prior <- brm(
  gi | se(sei) ~ 1 + (1 | article / experiment),
  data = dat,
  prior = prior_c,
  sample_prior = "only",
  save_pars = save_pars(all = TRUE),
  chains = 4,
  iter = 10000,
  cores = 4,
  control = list(adapt_delta = 0.99)
)
summary(res_prior)
# plot(res_prior)

@@

```{r, fig_forest}
# Forest plot inputs
level <- 0.05
ndraws <- 1000

@@

# Get draws from the Bayesian posterior distribution
draws <- epred_draws(res_brm, dat, ndraws = ndraws)
meta_draws <- spread_draws(res_brm, b_Intercept, ndraws = ndraws) %>%
  mutate(article = "Three-level model", experiment = "zzz") # For the last line

# Create forest plot
dat %>%
  # Compute frequentist confidence intervals
  mutate(
    ci_lb = gi - qnorm(level / 2, lower.tail = FALSE) * sqrt(vi),
    ci_ub = gi + qnorm(level / 2, lower.tail = FALSE) * sqrt(vi),
    fontface = "plain" # For study labels and confidence interval
  ) %>%
  # Add meta-analytic effect size + CI
  bind_rows(list(
    article = "Three-level model",
    group = "",
    experiment = "zzz", # Just so it's displayed as the last line
    gi = res_ml$b,
    ci_lb = res_ml$ci.lb,
    ci_ub = res_ml$ci.ub,
    fontface = "bold"
  )) %>%
  # Format effect size + CI for printing them on the plot
  mutate(ci_print = print_ci(gi, ci_lb, ci_ub)) %>%
  # Reverse order of experiments so they are plotted alphabetically
  mutate(experiment = forcats::fct_rev(factor(experiment))) %>%
  arrange(experiment) %>%
  # Show article labels only for the first experiment per article
  mutate(
    article = if_else(article == lead(article, default = ""), "", article)
  ) %>%
  # Prepare plot
  ggplot(aes(x = gi, y = experiment)) +
  geom_vline(xintercept = seq(-1, 2, 1), size = 0.8, color = "grey90") +
  geom_vline(xintercept = seq(-1.5, 2.5, 1), size = 0.4, color = "grey90") +
  # Add article and group labels as text to the left
  geom_text(aes(x = -5.6, label = article, fontface = fontface), hjust = 0) +
  geom_text(aes(x = -3.1, label = group, fontface = fontface), hjust = 0) +
  # Add experiment specific effect size + CI as text to the right
  geom_text(aes(x = 4.4, label = ci_print, fontface = fontface), hjust = 1) +
  # Add Bayesian highest posterior density interval in the background
  stat_interval(aes(x = .epred), draws, point_interval = "mean_hdi") +
  stat_interval(aes(x = b_Intercept), meta_draws, point_interval = "mean_hdi") +
  scale_color_grey(
    start = 0.9, end = 0.6,
    labels = as_mapper(~ scales::percent(as.numeric(.x)))
  ) +
  labs(color = "Bayesian\nHPDI") +
  # Add study specific effect size + CIs on top
  ggnewscale::new_scale_color() +
  geom_errorbar(aes(xmin = ci_lb, xmax = ci_ub, color = gender)) +
  geom_point(aes(size = ni, color = gender), shape = 15) + # Or (1 / vi)?
  scale_color_brewer(palette = "Dark2", na.translate = FALSE) +
  # Add meta-analytic effect size as a line
  annotate("segment", x = res_ml$b, xend = res_ml$b, y = 0.7, yend = 1.3) +
  annotate("errorbar", xmin = res_ml$ci.lb, xmax = res_ml$ci.ub, y = 1.0) +
  # Styling
  coord_cartesian(expand = FALSE) +
  scale_x_continuous(breaks = seq(-1.5, 2.5, 0.5)) +
  labs(
    x = expression("Hedges'" ~ italic("g")),
    size = "Sample\nsize",
    color = "Gender"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.title.x = element_text(hjust = 0.584),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    legend.title = element_text(hjust = 0.5),
    panel.grid = element_blank(),
    text = element_text(family = "Helvetica")
  )
